The project objectives are divided into primary and secondary objectives:

\subsection*{Primary Objectives}
\begin{enumerate}
    \item \textbf{Design a long-range wireless communication system} using the 433 MHz band with a minimum range of 500 meters.
    
    \item \textbf{Implement safety shutdown mechanism} that cuts power to robot motors within 100 ms of communication loss detection.
    
    \item \textbf{Develop manual override functionality} allowing the operator to control velocity (forward/reverse) and turn rate (left/right).
    
    \item \textbf{Design battery-powered remote} with minimum 8-hour operation time.
    
    \item \textbf{Design and build custom PCB} for both remote and robot units.
    
    \item \textbf{Design and 3D-print enclosures} for both devices.
\end{enumerate}

\subsection*{Secondary Objectives (Optional)}
\begin{enumerate}
    \item \textbf{Implement bidirectional communication} to display robot status (battery voltage, errors) on the remote.
    
    \item \textbf{Add low battery warning} to the remote display.
    
    \item \textbf{Implement signal strength indicator} (RSSI) on the remote.
\end{enumerate}

\subsection*{Success Criteria}
The project will be considered successful if:
\begin{itemize}
    \item Communication range exceeds 500 meters in open terrain
    \item Motor shutdown occurs within 100 ms of signal loss
    \item Remote operates for minimum 8 hours on a full battery charge
    \item Both units are fully functional prototypes with custom PCBs and enclosures
\end{itemize}
